---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OmopStudyBuilder <a href="https://github.com/oxford-pharmacoepi/OmopStudyBuilder"><img src="man/figures/image.jfif" align="right" height="138" alt="OmopStudyBuilder logo" /></a>

<!-- badges: start -->

[![R-CMD-check](https://github.com/oxford-pharmacoepi/OmopStudyBuilder/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/oxford-pharmacoepi/OmopStudyBuilder/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/OmopStudyBuilder)](https://CRAN.R-project.org/package=OmopStudyBuilder)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/oxford-pharmacoepi/OmopStudyBuilder/branch/main/graph/badge.svg)](https://app.codecov.io/gh/oxford-pharmacoepi/OmopStudyBuilder?branch=main)

<!-- badges: end -->

The **OmopStudyBuilder** package helps a study prepare for network studies using the OMOP Common Data Model (CDM). The package sets up an R project for your study with a default folder structure and template code typically required for network studies. This allows you to focus on the parts of the analysis that are specific to your study design.

In addition to project setup, the package reviews code and dependencies, supports renv locking for consistent package versions, and can build and run a Docker image for reproducible execution aligned with best practices.

The package is highly opinionated and designed to align with the OxInfer study code checklist. For further details, please refer to the formal documentation at:
[https://oxford-pharmacoepi.github.io/Oxinfer/onboarding/code_review.html](https://oxford-pharmacoepi.github.io/Oxinfer/onboarding/code_review.html)

# Installation

You can install the development version of the package from GitHub:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("oxford-pharmacoepi/OmopStudyBuilder")
```

## Installing Docker

This package supports building and running study code in Docker for reproducible execution.
Install Docker and confirm it is running before using the Docker-based workflow.

**General checks (all operating systems)**

- Install Docker.
- Start Docker (e.g., Docker Desktop).
- Verify Docker is available:
  - `docker --version`
  - `docker info`

**macOS**

- Install Docker Desktop for Mac: https://docs.docker.com/desktop/setup/install/mac-install/
- Open Docker Desktop and wait until it reports “Docker Desktop is running”.

**Windows**

- Install Docker Desktop for Windows (WSL 2 recommended): https://docs.docker.com/desktop/setup/install/windows-install/
- Ensure WSL 2 is enabled (Docker Desktop installer will guide you).
- Open Docker Desktop and wait until it is running.

**Linux**

- Install Docker Engine using your distribution instructions: https://docs.docker.com/engine/install/
- Start/enable the Docker service (varies by distro).
- If you want to run Docker without `sudo`, follow: https://docs.docker.com/engine/install/linux-postinstall/

# Example Usage

To illustrate how OmopStudyBuilder works, start by creating the study folder and reviewing what it contains:

```{r, eval = FALSE}
library(OmopStudyBuilder)
createStudy(here::here("SampleStudy"))

reviewStudyCode(here::here("SampleStudy", "study_code"))
reviewStudyDependencies(here::here("SampleStudy", "study_code"))
```

Lock package versions with renv so everyone runs the same environment, then build the study image from the study folder.

```{r, eval = FALSE}
renv::init(here::here("SampleStudy", "study_code"))
install.packages(c("dplyr", "CDMConnector", "IncidencePrevalence"))
renv::snapshot(here::here("SampleStudy", "study_code"))

buildStudy(path = here::here("SampleStudy", "study_code"))
```

Run the study interactively in RStudio Server or as an automated script. If you use a `.env` file for credentials, pass `env_file = ".env"`.

```{r, eval = FALSE}
runRStudio()
runStudy()
```

Use optional inputs only if you need them:

```{r, eval = FALSE}
runStudy(
  image_name = "omop-study-study-code",
  data_path = "path/to/data",
  results_path = "./results"
)
```

To distribute the study, share the **study folder** created by `createStudy()` (including `study_code/` and `renv.lock`). Partners can build and run using the same commands.

```{r, eval = FALSE}
install.packages("OmopStudyBuilder")
library(OmopStudyBuilder)

buildStudy(path = here::here("SampleStudy", "study_code"))
runStudy()
runRStudio()
stopStudy()
```
