---
output:
  github_document:
    html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OmopStudyBuilder <img src="man/figures/image.jfif" align="right" height="138" alt="OmopStudyBuilder logo" />

<!-- badges: start -->

[![R-CMD-check](https://github.com/oxford-pharmacoepi/OmopStudyBuilder/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/oxford-pharmacoepi/OmopStudyBuilder/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/OmopStudyBuilder)](https://CRAN.R-project.org/package=OmopStudyBuilder)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/oxford-pharmacoepi/OmopStudyBuilder/branch/main/graph/badge.svg)](https://app.codecov.io/gh/oxford-pharmacoepi/OmopStudyBuilder?branch=main)

<!-- badges: end -->

The **OmopStudyBuilder** package helps a study prepare for network studies using the OMOP Common Data Model (CDM). The package sets up an R project for your study with a default folder structure and template code typically required for network studies. This allows you to focus on the parts of the analysis that are specific to your study design.

In addition to project setup, the package reviews code and dependencies, supports renv locking for consistent package versions, and can build and run a Docker image for reproducible execution aligned with best practices.

The package is highly opinionated and designed to align with the OxInfer study code checklist. For further details, please refer to the [documentation](https://oxford-pharmacoepi.github.io/Oxinfer/onboarding/code_review.html).

# Installation

You can install the development version of the package from GitHub:

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("oxford-pharmacoepi/OmopStudyBuilder")
```

# Example Usage

To illustrate how OmopStudyBuilder works, start by creating the study folder and reviewing what it contains:

```{r, eval = FALSE}
library(OmopStudyBuilder)
createStudy(here::here("SampleStudy"))

reviewStudyCode(here::here("SampleStudy/study_code"))
reviewStudyDependencies(here::here("SampleStudy/study_code"))
```

Lock package versions with renv so everyone runs the same environment, then build the study image from the study folder.

```{r, eval = FALSE}
renv::init(here::here("SampleStudy/study_code"))
install.packages(c("dplyr", "CDMConnector", "IncidencePrevalence"))
renv::snapshot(here::here("SampleStudy/study_code"))

buildStudy(path = here::here("SampleStudy/study_code"))
```

Run the study interactively in RStudio Server or as an automated script. If you use a `.env` file for credentials, pass `env_file = ".env"`.

```{r, eval = FALSE}
runRStudio()
runStudy()
```

Use optional inputs only if you need them:

```{r, eval = FALSE}
runStudy(
  image_name = "omop-study-study-code",
  data_path = "path/to/data",
  results_path = "./results"
)
```

To distribute the study, share the **study folder** created by `createStudy()` (including `study_code/` and `renv.lock`). Partners can build and run using the same commands.

```{r, eval = FALSE}
install.packages("OmopStudyBuilder")
library(OmopStudyBuilder)

buildStudy(path = here::here("SampleStudy/study_code"))
runStudy()
runRStudio()
stopStudy()
```

To push a built image to Docker Hub, use the helper and enter your credentials when prompted (the tag defaults to `latest`, and the image name defaults to the current folder name):

```{r, eval = FALSE}
pushStudyImage(
  repo = "yourname/omop-study-study-code"
)
```
