---
title: "createStudy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{createStudy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(OmopStudyBuilder)
library(here)
```

## Introduction

`createStudy()` is the starting point for building a study project with the
**OMOP Common Data Model (CDM)** using the `OmopStudyBuilder` package.

It generates a fully structured project folder containing:

* Templates for study diagnostics
* Templates for analysis code
* Shiny apps for diagnostics and reporting
* A project README
* A reproducible, opinionated layout aligned with [OxInfer](https://oxford-pharmacoepi.github.io/Oxinfer/onboarding/code_review.html) conventions

This allows you to immediately focus on writing study-specific code instead of
manually assembling the surrounding infrastructure.

In this vignette you will learn how to:

* Create a new study project using `createStudy()`
* Explore the folder structure that is generated
* Control whether diagnostics and/or study analysis templates are included
* Understand next steps after project creation

---

## Basic usage

The simplest way to use `createStudy()` is to provide a directory path.
If the directory:

* **Does not exist**: it will be created
* **Exists but is empty**: it will be used
* **Exists and contains files**: an error will be thrown

To avoid modifying your real project directories, this vignette uses a temporary
folder.

```{r basic-usage, message=FALSE}
# Create a temporary root for this vignette
study_root <- file.path(tempdir(), "SampleStudy")

createStudy(study_root)

# Show top-level contents created by createStudy()
list.files(study_root)
```

You should see something like:

* `diagnostics_code/` – diagnostic scripts and helpers
* `diagnostics_shiny/` – skeleton diagnostics Shiny app
* `study_code/` – scripts for running the main analyses
* `study_shiny/` – skeleton reporting Shiny app
* `README.md` – overview of the study project

To inspect the full structure:

```{r list-recursive}
list.files(study_root, recursive = TRUE)
```

---

## Diagnostics-only and study-only setups

You can control which parts of the template are created with the
`diagnostics` and `study` arguments.

### Diagnostics-only project

```{r diagnostics-only, message=FALSE}
diag_root <- file.path(tempdir(), "DiagnosticsOnly")

createStudy(
  directory   = diag_root,
  diagnostics = TRUE,
  study       = FALSE
)

list.files(diag_root)
```

### Study-only project (no diagnostics templates)

```{r study-only, message=FALSE}
study_only_root <- file.path(tempdir(), "StudyOnly")

createStudy(
  directory   = study_only_root,
  diagnostics = FALSE,
  study       = TRUE
)

list.files(study_only_root)
```

---

## What to do after `createStudy()`

After generating the project structure, you will mainly work with these folders and files:

- `study_code/`  
  - `StudyCode.Rproj`: open this in RStudio to work on the main analysis code. It sets the working directory so all relative paths behave correctly.  
  - `code_to_run.R`: where you add database connection details, schemas, and global settings
  - `cohorts/instantiate_cohorts.R`: defines the study cohorts using your codelists and cohort-building functions. You customise this to match your phenotype definitions.  
  - `run_study.R`: orchestrates the analysis steps (cohort creation, summaries, analyses, export). 

- `diagnostics_code/`  
  - `diagnostics.Rproj`: open this project to work on diagnostics. It keeps diagnostics code separate from the main analysis project.  
  - `code_to_run.R`: similar to the study version, but dedicated to diagnostics. You set connection details and schemas used for running phenotype diagnostics.  
  - `cohorts/instantiate_cohorts.R`: defines cohorts for diagnostic checking, often mirroring or simplifying the main study cohorts.  
  - `run_study.R`: runs the PhenotypeDiagnostics workflow and exports diagnostic results. You might tweak options, but the overall flow is usually left intact.

- `study_shiny/`  
  - For a shiny app to explore study results. You extend this with plots, tables, and filters built on top of your exported outputs.

- `diagnostics_shiny/`  
  - Contains a minimal Shiny app for exploring diagnostic outputs. Use it to review cohort characteristics and diagnostic results interactively across data partners.

---

## Building and Distributing Your Study

Once your study code is ready, `OmopStudyBuilder` provides functions to package it into a Docker image and distribute it to data partners.

### Building the Docker Image

```{r build-docker, eval=FALSE}
# Build a Docker image containing your study code
buildStudy("MyStudy/study_code")
#> Building Docker image: omop-study-study-code
#> This may take 5-15 minutes on first build...
#> ✔ Image built successfully: omop-study-study-code
```

The image contains:
- Your exact R version
- All package dependencies (from renv.lock)
- Your study code
- Database drivers and system dependencies

### Distribution Options

#### Option 1: Docker Hub (Online)

Push to Docker Hub for easy sharing with connected partners:

```{r push-image, eval=FALSE}
pushImage("omop-study-study-code", "yourusername/myomopstudy:v1.0")
#> ✔ Image pushed to Docker Hub!
```

Data partners pull and run:

```{r pull-run, eval=FALSE}
library(OmopStudyBuilder)
pullImage("yourusername/myomopstudy:v1.0")
runRStudio(
  image_name = "yourusername/myomopstudy:v1.0",
  data_path = "C:/my_omop_data"
)
```

#### Option 2: Tar File (Offline)

Export to a file for secure/offline distribution:

```{r export-image, eval=FALSE}
exportImage("omop-study-study-code", "myomopstudy-v1.0.tar")
#> ✔ Image exported: myomopstudy-v1.0.tar (1.45 GB)
```

Data partners load and run:

```{r load-run, eval=FALSE}
library(OmopStudyBuilder)
loadImage("myomopstudy-v1.0.tar")
runRStudio(data_path = "C:/my_omop_data")
```

### Interactive vs Automated Execution

**RStudio Server** (recommended for non-technical partners):

```{r rstudio-run, eval=FALSE}
runRStudio(data_path = "path/to/data")
#> ✔ RStudio Server started!
#> ! Open browser: http://localhost:8787
#> ! Username: rstudio
#> ! Password: study1234
```

Partners edit credentials in `code_to_run.R` and run the study interactively.

**Automated execution** (for programmatic workflows):

```{r auto-run, eval=FALSE}
runStudy(
  image_name = "omop-study-study-code",
  data_path = "path/to/data",
  output_path = "./results"
)
```

See `inst/DATA_PARTNER_GUIDE.md` for detailed instructions to share with data partners.

---
